An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-21 VLDB2026/main.tex

	\textit{\textbf{Multi-Attribute Building and Multi-Attribute Query.}}  
	Compared to single-attribute filtering, multi-attribute joint filtering better reflects real-world scenarios. To evaluate algorithm performance under this setting, we construct the index using 3 uniformly distributed attributes and apply filtering on all 3 attributes simultaneously during queries.
	%	\textcolor{red}{As Filtered-DiskANN does not support the AND rule used in our multi-attribute queries, it is excluded.}
	%	Compared to single-attribute filtering, multi-attribute joint filtering better reflects real-world application scenarios. To evaluate algorithm performance under such scenario, we construct the index using 3 uniformly distributed attributes and apply filtering conditions on all 3 attributes simultaneously during the query process.

@@

	As shown in Figure~\ref{fig:exp_6_1}, all methods achieve their highest QPS and recall on Audio dataset. Because Audio has small scale, low dimensionality, and low LID. In contrast, all methods exhibit the lowest QPS and recall on the GIST1M and GloVe dataset besides the OOD dataset Text2Image. Because these two datasets have the largest scale and highest LID.

@@

	\textit{\textbf{\textcolor{blue}{Complex Queries.}}}

@@

	\textit{\textbf{\textcolor{blue}{Complex Queries.}}}

>>>>

	\textit{\textbf{Multi-Attribute Building and Multi-Attribute Query.}}  
	Compared to single-attribute filtering, multi-attribute joint filtering better reflects real-world scenarios. To evaluate algorithm performance under this setting, we construct the index using 3 uniformly distributed attributes and apply filtering on all 3 attributes simultaneously during queries. \textcolor{orange}{Note that FilterDiskANN only supports Boolean OR logic and does not support Boolean AND. Therefore, its two Vamana-based methods are excluded from the multi-attribute construction and search experiments.}
	%	\textcolor{red}{As Filtered-DiskANN does not support the AND rule used in our multi-attribute queries, it is excluded.}
	%	Compared to single-attribute filtering, multi-attribute joint filtering better reflects real-world application scenarios. To evaluate algorithm performance under such scenario, we construct the index using 3 uniformly distributed attributes and apply filtering conditions on all 3 attributes simultaneously during the query process.

@@

%ykw: 这段话看着有问题
	As shown in Figure~\ref{fig:exp_6_1}, all methods achieve their highest QPS and recall on Audio dataset. Because Audio has small scale, low dimensionality, and low LID. In contrast, all methods exhibit the lowest QPS and recall on the GIST1M and GloVe dataset besides the OOD dataset Text2Image. Because these two datasets have the largest scale and highest LID.

@@

	\textit{\textbf{\textcolor{blue}{Large-scale Datasets}}}

	\textit{\textbf{\textcolor{blue}{Multi-modal Queries}}}

	\textcolor{blue}{
	We conducted experiments on the text2image-mix dataset. This dataset contains both image and text vectors. It is a more complex multimodal dataset. As shown in Figure~\ref{fig:TrinityExtension——Attribution}, although the queries are more complex, UNG and NHQ still perform the best. CAPS is the next best.
	Since the dataset mixes two modalities, the ground truth of the query set usually corresponds to the same modality. The nearest neighbors in the index are often from the same modality because they are closer in distance. Therefore, the search results on this dataset are similar to those in Figure~\ref{fig:exp_1_1_SingleLabel_1thread}. UNG and NHQ perform the best. CAPS clusters based on vector similarity. It actually clusters the vectors into the same modality. Then it uses attribute filtering to divide the data further. This allows CAPS to quickly find nearest neighbors that meet the filter conditions.
	In contrast, the text2image dataset shown in Figure~\ref{fig:exp_6_1} is different. The queries and the dataset are from different modalities. The queries are far from the dataset. The ground truths are also scattered and not clustered together. To find the ground truths, the algorithm often needs to traverse more points. This is a greater challenge for the algorithms.
%	We conducted extensive scalability experiments on the Text2Image-Mix dataset. This dataset contains both image and text vectors. It represents a more complex multimodal query scenario. As shown in Figure 12, the UNG and NHQ algorithms still achieve the best performance, followed closely by CAPS.
%	In the Text2Image-Mix dataset, the ground truth of the query vector is usually the corresponding vector in the same modality. Since the neighbor relationships in the graph are often based on vector distance, vectors of similar modalities naturally form closer connections in the graph. Therefore, during the search process, the algorithm tends to traverse efficiently within the same modality. This makes the search mechanism similar to our experiments on single-modality datasets, thus maintaining high performance.
%	The good performance of the CAPS algorithm is due to its unique hierarchical space partitioning strategy. It first clusters vectors based on similarity. In multimodal data, this clustering naturally groups vectors of similar modalities into the same cluster. Then, CAPS further partitions these clusters using attributes. This allows it to quickly find the nearest neighbors that meet the filtering conditions.
%	It is worth noting that, compared to the OOD dataset discussed in the "Dataset Impact" section, the Text2Image-Mix dataset also involves modality differences. However, its complexity lies in modality mixing. The OOD dataset, on the other hand, has a modality gap. This often requires traversing a broader and more sparse search space to find the ground truth. It poses a more severe challenge to the algorithm.
}

@@

	\textit{\textbf{\textcolor{blue}{Multi-modal Queries}}}
	\textcolor{blue}{
	Similar to attribute filtering, the performance of multimodal queries with range filtering shown in Figure~\ref{fig:MultiModel range} is also consistent with the single-modality results in Figure~\ref{fig:exp_8_2}. As shown in Figure~\ref{fig:exp_8_3}, the algorithms still perform the worst on multimodal datasets like Text2Image. This is because these range query algorithms are almost graph-based. Graph algorithms need more iterations to find true neighbors in OOD datasets.}
